---
title: "DS 520 Lab 1"
output: pdf_document
---

Task 1:

Load "mtcars" data and create a new dataset having “mpg”, “cyl”,“disp”,“hp”, “wt”, and “am”
View it and the dataframe

```{r}
mtcars
newdataset <- mtcars[,c("mpg","cyl","disp","hp","wt","am")]
View(newdataset)
str(mtcars)
```


Task 2:

For the new dataset, create a new categorical variable “mpg_group” where,
If miles per gallon less than 12, then “low”;
If miles per gallon less than 20 and bigger or equal to 12, then “medium”;
Otherwise, “high”.
and add it to your new dataset.

```{r}
attach(newdataset)
newdataset$mpg_group[mpg < 12] <- "low"
newdataset$mpg_group[mpg >= 12 &  mpg < 20] <- "medium"
newdataset$mpg_group[mpg >= 20] <- "high"
detach(newdataset)
```


Task 3:

For each level of “mpg_group”, attach the average of miles per gallon (“mpg”), displacement (“disp”), and weight (“wt”).

```{r}
attach(newdataset)
newdataset_1 <- newdataset[,c("mpg","disp","wt","mpg_group")]
agg_data <- aggregate(newdataset_1, by=list(mpg_group),FUN = mean, na.rm=TRUE)
```

newdataset_1 created because newdataset has other variables needed later so cannot set it to newdataset again


Task 4:

Merge your dataset from task 3 back to your new data by “mpg_group”.

```{r}
merge_data <- merge(agg_data,newdataset,by.x=c("Group.1"),by.y=c("mpg_group"),ALL=TRUE)
```

mpg_group turned into Group.1 in aggregate function because it couldn't take mean of words


Task 5:

Create a subset of your merged dataset where number of cylinders less than 8 and transmission is automatic. Finally, clean up the dataset you just created.

```{r}
subset_data <- subset(merge_data, cyl < 8 & am == 1)
subset_data[,-5]
```

removed the mpg_group which contained NA